QPLIB: A Library of Quadratic Programming Instances

### 2017-01-20 QPLIB-3.tex
say gams performance libraries, fix capitalization of GAMS
- GAMS has a number of libraries, so "GAMS library" is very unspecific

In addition to spontaneous contribution we analysed the other generic
libraries of instances available  on internet and containing QP
instances. Namely, the libraries from which we gathered instances are
%
\begin{itemize}
 \item the {\tt BARON} library
 \url{http://www.minlp.com/nlp-and-minlp-test-problems};
%
\item the {\tt CUTEst} library
 \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki};
%
\item the {\tt GAMS} library
 \url{http://www.gamsworld.org/performance/performlib.htm};
%
\item the {\tt MacMINLP} library
 \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP};
%
\item the {\tt Meszaros} library
 \url{http://www.doc.ic.ac.uk/~im/00README.QP};
%
\item the {\tt MINLP} library
 \url{http://www.gamsworld.org/minlp/minlplib.htm};
%
\item the {\tt POLIP} library
 \url{http://polip.zib.de/pipformat.php}.
\end{itemize}

@@

To evaluate the instances we decided, for practical reasons, to use
Gams as common platform for all the experiments involving commercial and
non-commercial solvers. For this reason, we decided to translate all
instances into the Gams format (\texttt{.gms}). 
% In a preliminary phase, all the instances received were divided
%according to their format and subsequently translated.
%In \S~\ref{subsec:tools}
%%~\ref{subsec:CP_convex}
%the tools used
%to translate an instance from a given format to the \texttt{.gms} format
%%to translate an instance from and to a given format to the
%\texttt{.gms} format
%are described in more detail.\\
%
In addition, we have introduced a new, specific \texttt{.qplib} format. This
new format is capable of describing all the instances of the library in a sparse form.
In comparison to a more \emph{high level} format like \texttt{.gms}, the new
format presents two advantages: it is easier to read by a self-made parser and
it produces smaller files. See the Appendix~\label{sec:format} for more details.

@@

During the development of the library, a discussion ensued about what
the expected gols that we wanted to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item represent as much as possible all the different categories of QP
       problems;
 %
 \item gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item include for each of the categories a set of well-diversified
       instances;
 %
 \item obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve the aforementioned goals, we performed the following two
filters, applied in cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S \ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in {\tt Gams} (cf.~Table \ref{t:solvers})
       capable of solving it, whose number depends on the category of the
       instance under consideration. We then filtered according to a relative
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives. {\bf Comment: we may want to be a tad more specific
       about what ``features'' exactly meant there}. Finally, in order to only
       keep computationally challenging instances we ran a general purpose solver
       able to tackle all QP categories with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded. {\bf Comment: this might actually be best characterized as
       a third filter.}
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $251$ discrete instances and
$116$ continuous instances.

>>>>

In addition to spontaneous contribution we analysed the other generic
libraries of instances available  on internet and containing QP
instances. Namely, the libraries from which we gathered instances are
%
\begin{itemize}
 \item the {\tt BARON} library
 \url{http://www.minlp.com/nlp-and-minlp-test-problems};
%
\item the {\tt CUTEst} library
 \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki};
%
\item the {\tt GAMS Performance} libraries
 \url{http://www.gamsworld.org/performance/performlib.htm};
%
\item the {\tt MacMINLP} library
 \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP};
%
\item the {\tt Meszaros} library
 \url{http://www.doc.ic.ac.uk/~im/00README.QP};
%
\item the {\tt MINLP} library
 \url{http://www.gamsworld.org/minlp/minlplib.htm};
%
\item the {\tt POLIP} library
 \url{http://polip.zib.de/pipformat.php}.
\end{itemize}

@@

To evaluate the instances we decided, for practical reasons, to use
GAMS as common platform for all the experiments involving commercial and
non-commercial solvers. For this reason, we decided to translate all
instances into the GAMS format (\texttt{.gms}). 
% In a preliminary phase, all the instances received were divided
%according to their format and subsequently translated.
%In \S~\ref{subsec:tools}
%%~\ref{subsec:CP_convex}
%the tools used
%to translate an instance from a given format to the \texttt{.gms} format
%%to translate an instance from and to a given format to the
%\texttt{.gms} format
%are described in more detail.\\
%
In addition, we have introduced a new, specific \texttt{.qplib} format. This
new format is capable of describing all the instances of the library in a sparse form.
In comparison to a more \emph{high level} format like \texttt{.gms}, the new
format presents two advantages: it is easier to read by a self-made parser and
it produces smaller files. See the Appendix~\label{sec:format} for more details.

@@

During the development of the library, a discussion ensued about what
the expected gols that we wanted to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item represent as much as possible all the different categories of QP
       problems;
 %
 \item gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item include for each of the categories a set of well-diversified
       instances;
 %
 \item obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve the aforementioned goals, we performed the following two
filters, applied in cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S \ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in {\tt GAMS} (cf.~Table \ref{t:solvers})
       capable of solving it, whose number depends on the category of the
       instance under consideration. We then filtered according to a relative
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives. {\bf Comment: we may want to be a tad more specific
       about what ``features'' exactly meant there}. Finally, in order to only
       keep computationally challenging instances we ran a general purpose solver
       able to tackle all QP categories with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded. {\bf Comment: this might actually be best characterized as
       a third filter.}
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $251$ discrete instances and
$116$ continuous instances.
